---
title: "Week 1 Science and Cooking questions"
output: html_notebook
---

So far I have only included the quantitative questions here below:

Question:
molar mass of saccharin: C7H5NO3S

```{r}
c = 12
h = 1
N = 14
O = 16
S = 32

7*c + 5*h + N + 3*O + S

```

Nutrition Label - Moles Calculation
2 points possible (ungraded)
Now let's look at the nutrition label again.

Nutrition Facts	
Serving size	55 grams
Amount per serving	
Calories	230
% Daily Value
Total Fat 8 g 12%
Saturated Fat 1 g 5%
Cholesterol 0 mg 0%
Sodium 160 mg 7%
Carbohydrates 37 g 12%
Dietary Fiber 4 g 16%
Sugars 1 g	
Protein 3 g

A. If you assume that the molecular weight of sugar is 342g/mol, how many moles of sugar are there in one serving? Enter your answer in the box below, and do not include the units

```{r}
1/342
```


B. If you assume that the molecular weight of sodium is 23g/mol, how many moles of sodium are there in one serving?Enter your answer in the box below, and do not include the units.

```{r}
160*10^-3/23
```

If the concentration of water molecules in neutral water is 55 moles/L, and the concentration of hydrogen ions is determined by the pH, what fraction of the molecules in neutral water are hydrogen ions? (Enter your answer as a percentage without a % sign.)

ph neutral water = 7

```{r}
ph = 7

M_H = 10^(-1*ph)

(M_H/55)*100
```

When one molecule of lye (NaOH) dissolves, it can “neutralize” one proton. A particular recipe for pretzels suggests making a 10% solution by weight of lye in water for pretzel dipping. This is the equivalent of saying that 100g of sodium hydroxide is dissolved in 1L of water. Given a molecular weight of 40g/mol for sodium hydroxide, how many moles of acid can be neutralized by the lye in a 1L pretzel bath? Enter the number of moles, without including units.

```{r}
M_lye = 100/40
#1 M_lye = 1 M_Na + 1 M_OH
M_lye

```





